`Total-evidence' Bayesian estimation of phylogeny, divergence times and fossil ages

### 2016-04-06 response-to-reviewers.tex

\documentclass[11pt]{article}

\usepackage{xcolor}

\usepackage[numbers]{natbib}

\newcommand{\response}[1]{\medskip{}\textcolor{blue}{{Response: #1}}\medskip{}}
\newcommand{\respquote}[1]{\medskip{}\textcolor{blue}{``{\it #1}''}\medskip{}}

@@

p.1. Abstract: change ``phylogenetic age'' by ``phylogenetic estimates of fossil age''. Please change throughout the ms.

\response{We have replaced all instances of ``phylogenetic age'' with variants of ``phylogenetic estimates of fossil age''.}

p.4. Data set descriptions: For each dataset please add a small paragraph providing additional details, i.e. total number of characters, amount of missing data, fossil age ranges (say, penguin fossils range from xx Mya to yy Mya), and character state space (i.e. characters ranged from binary to k=x).

\response{We added the following paragraph:}

\respquote{We used a data set originally published by \cite{ksepka2012} consisting of morphological data from fossil and living penguin 
species. 
We used the same subset of this original data as in \cite{gavryushkina2015bayesian}. 
Additionally we discarded the molecular sequence data from the living species. 
The resulting morphological data matrix we used contains 36 fossil species, 19 extant species and the morphological matrix contains 202 characters. 
The morphological characters ranged from binary to k=7. 
The majority of these (\textgreater 95\%) have fewer than four states and 48 of the binary characters were encoded as presence/absence. 
See \cite{gavryushkina2015bayesian} for further details.}

### 2016-04-06 fossilDating-revision.tex
phylogenetic analyses

\newcommand{\Mstrict}{{M1}}
\newcommand{\Mrelaxed}{{M8}}

@@

%TS: check:
For each of the 36 penguin fossils in turn we performed a separate Bayesian phylogenetic analysis in which the focal fossil's palaeontological age constraints were replaced by 
the fossilized birth-death process prior as described below.  
For the choices of morphological substitution model and tree prior we followed \cite{gavryushkina2015bayesian} and considered two of her modelling combinations. 
The first was \Mstrict{} in which a strict clock and the simplest model of morphological substitution was assumed \cite{Lewis2001}.
We also employed \Mrelaxed{} from \cite{gavryushkina2015bayesian} which involved a more complex model assuming a relaxed clock model \cite{Drummond2006}, partitioning of the sites into groups of equal state count and an additional parameter for gamma-distributed rate variation across sites \cite{yang:1994ma}.

@@

>>>>

\newcommand{\Mstrict}{{Mk-1}}
\newcommand{\Mrelaxed}{{Mk-8}}

@@

We used a data set originally published by \cite{ksepka2012} consisting of morphological data from fossil and living penguin 
species. 
We used the same subset of this original data as in \cite{gavryushkina2015bayesian}. 
Additionally we discarded the molecular sequence data from the living species. 
The resulting morphological data matrix we used contains 36 fossil species, 19 extant species and the morphological matrix contains 202 characters. 
The morphological characters ranged from binary to k=7. 
The majority of these (\textgreater 95\%) have fewer than four states and 48 of the binary characters were encoded as presence/absence. 
See \cite{gavryushkina2015bayesian} for further details.

For each of the 36 penguin fossils in turn we performed a separate Bayesian phylogenetic analysis in which the focal fossil's palaeontological age constraints were replaced by 
the fossilized birth-death process prior as described below.  
For the choices of morphological substitution model and tree prior we followed \cite{gavryushkina2015bayesian} and considered two of her modelling combinations. 
The first was \Mstrict{} in which a strict clock and the simplest model of morphological substitution was assumed \cite{Lewis2001}.
We also employed \Mrelaxed{} from \cite{gavryushkina2015bayesian} which involved a more complex model assuming a relaxed clock model \cite{Drummond2006}, partitioning of the sites into groups of equal state count and an additional parameter for gamma-distributed rate variation across sites \cite{yang:1994ma}.

@@

\subsection*{Phylogenetic analyses}

Following \cite{gavryushkina2015bayesian}, we apply the Lewis Mk model 
\citep{Lewis2001} for discrete morphological character evolution, which assumes a character can take $k$ states and the transition rates from one state to another are equal for all states. As did \cite{gavryushkina2015bayesian}, we treat 34 characters that were ordered in \citep{ksepka2012} as unordered.

We applied two phylogenetic models to each data set: Mk-1 and Mk-8. Mk-1 assumed a strict morphological clock and no gamma-distributed rate heterogeneity among sites. Model Mk-8 partitioned the alignment into six partitions, with the $i$'th partition containing all characters that had $k=i+1$ character states across the sampled taxa.
